---
title: "Resistance test in liquid"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE}
library(tidyverse)
library(broom)
library(cowplot)
library(here)
library(caTools)
```

# Reduction of Bacterial Growth 
Based on method described by Poullain et al. (2008). The evolution of specificity in evolving and coevolving antagonistic interactions between a bacteria and its phage. Evolution: International Journal of Organic Evolution, 62(1), 1-11.

>  The reduction in bacterial absorbance “RBG” induced by phage i on bacterial colony j, was calculated as RBGij=[abs630(t= 20) − abs630(t= 0)]ij/[abs630(t= 20) − abs630(t= 0)]controlj. Each cross‐infection was performed in triplicate at the same time. Thus only data corresponding to RBGij < 0.95 were considered in the analyses (positive infection). Data are presented as 1 – RBGi for more clarity.  

## load data  
Reads were every 5 minutes, averaging 3 time point window

```{r, message=FALSE}
input <- read_csv(here("data","shake_plate.csv"))
#time in hours
input$Time <- 24*input$Time
    
    #############
    #average OD #
    #############
input.avg <- 
input%>%
  map_dfc(function(x) runmean(x,k=3))


#trnaform to long format
d.avg <- gather(input.avg, key="well", value = "OD600", colnames(input)[-c(1:2)])
d <- gather(input, key="well", value = "OD600", colnames(input)[-c(1:2)])

#add metadata
meta <- read_csv(here("data","RBG_meta.csv"))

d.avg <-merge(d.avg, meta)
d <-merge(d, meta)

d.avg%>%
  # filter(Time<8)%>%
  # filter(dilution!=1e-7 & dilution!=1e-6)%>%
  group_by(well)%>%
  ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage))+
    facet_grid(row~col)+
  theme_cowplot()+
  ggtitle("Average")

d%>%
  # filter(Time<8)%>%
  # filter(dilution!=1e-7 & dilution!=1e-6)%>%
  group_by(well)%>%
  ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage))+
    facet_grid(row~col)+
  theme_cowplot()+
  ggtitle("Raw")

```

Averaging helps smooth things, as expected. 

Let's first look at the ansector, where all should be suceptible

```{r}
d.avg%>%
  filter(str_detect(host, "ANC"))%>%
  group_by(well)%>%
  ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border() +
  background_grid(major = "x",minor = "x")

```

There is growth of resistant culture at 5<T<10

```{r}
d.avg%>%
  filter(Time<10)%>%
  filter(str_detect(host, "ANC"))%>%
  group_by(well)%>%
  ggplot(aes(x=Time, y=OD600))+
  geom_vline(xintercept = 6, color="red")+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border() +
  background_grid(major = "x",minor = "x")+
  scale_x_continuous(breaks = 1:10)
```

6 hours is the shift.
```{r}
Tmax <- 6
```

```{r}
d.avg%>%
  filter(Time<Tmax)%>%
  # filter(str_detect(host, "ANC"))%>%
  group_by(well)%>%
  ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border() 
```

# Calculate RBG

```{r}
rbg <-  
d.avg%>%
  filter(Time<Tmax)%>%
  # get first and lat time point within 0:Tmax
    group_by(well) %>%
  arrange(Time) %>%
  filter(row_number()==1 | row_number()==n())%>%
  mutate(Time=ifelse(Time<1,"start","end"))%>%
  # pivot to join reads of me well
  pivot_wider(id_cols=c("well", "host", "phage"),names_from = Time, values_from=OD600)%>%
  mutate(dif =end-start)%>%
  # divide dif of each phage by no phage
  ungroup()%>%
  group_by(host) %>%
   mutate(rbg.v = dif / dif[phage == 'noPhage'])%>%
  ungroup()%>%
  filter(phage!="noPhage")%>%
  filter(!str_detect(host, "blank"))%>%
    mutate(phage.time= case_when(phage=="futuure" ~ 1,
                               phage=="current" ~ 0,
                               phage=="past" ~ -1))%>%
  mutate(evo.host= if_else(str_detect(host,"ANC"), "anc","evol"))
  
rbg %>%   
ggplot(aes(x=1, y=rbg.v))+
  geom_point(aes(color=phage))+
  facet_grid(.~host)+
  theme_cowplot()+panel_border()+
  xlab("resistance")
  
  # summarize over population
  
```

# show time shift curves
```{r}
rbg.mean <- rbg%>%
  filter(!str_detect(host,"ANC"))%>%
  mutate(rbg.v=if_else(rbg.v>1,1,rbg.v))%>%
  group_by(phage.time)%>%
  summarise(rbg.avg=mean(rbg.v),rbg.sd=sd(rbg.v), .groups="drop" )
  

rbg%>%
  filter(evo.host=="evol")%>%
  mutate(rbg.v=if_else(rbg.v>1,1,rbg.v))%>%
  ggplot(aes(x=phage.time, y=rbg.v))+
  geom_line(aes(group=host,color=host), alpha=0.5, size=1)+
  geom_line(data=rbg.mean,aes(y=rbg.avg), size=2)+
  geom_pointrange(data=rbg.mean,
                  aes(y=rbg.avg,ymin=rbg.avg-rbg.sd,ymax=rbg.avg+rbg.sd),
                  size=1)+
  theme_cowplot()+
  scale_x_continuous(breaks=c(-1,0,1))+
  scale_color_grey()+
  theme(legend.position = "none")
```

# static plate data
```{r}
d.static <- read_csv(here("data", "static_plate.csv"))%>%
    #long format
  pivot_longer(-Time,names_to="well", values_to="OD600")%>%
    # rearrnge column order
  select(well,Time, OD600)%>%
  full_join(., meta, by="well")
```

## Looking at the controls (compared with shaken)
```{r}
pA <- 
d.avg%>%
  filter(str_detect(host, "ANC")|str_detect(host, "blank"))%>%
  # filter(Time<Tmax)%>%
  # get first and lat time point within 0:Tmax
    group_by(well) %>%
  arrange(Time) %>%
  filter(row_number()==1 | row_number()==n())%>%
  mutate(Time=ifelse(Time<1,0,2))%>%
      ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border()+
  theme(legend.position = "bottom",
        axis.text.x = element_blank())+
  ylim(0,1.25)+
  ggtitle("Shaken")

pA2 <- 
d.avg%>%
  filter(str_detect(host, "ANC")|str_detect(host, "blank"))%>%
  filter(Time<Tmax)%>%
  # get first and lat time point within 0:Tmax
    group_by(well) %>%
  arrange(Time) %>%
  filter(row_number()==1 | row_number()==n())%>%
  mutate(Time=ifelse(Time<1,0,2))%>%
      ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border()+
  theme(legend.position = "bottom",
        axis.text.x = element_blank())+
  ylim(0,1.25)+
  ggtitle("Shaken")

pB <- d.static%>%
   mutate(Time = case_when(Time=="T0" ~ 0,
                                Time=="Tend" ~ 1,
                                Time=="Tend.vortex" ~ 2))%>%
  filter(str_detect(host, "ANC")|str_detect(host, "blank"))%>%
  filter(Time!=1)%>%
    ggplot(aes(x=Time, y=OD600))+
    geom_line(aes(color=phage), size=1)+
    facet_wrap(~host)+
  theme_cowplot()+
    panel_border()+
  theme(legend.position = "bottom",
        axis.text.x = element_blank())+
  ylim(0,1.25)+
  ggtitle("Static")

plot_grid(pA, pB)+
  ggsave2(here("plots", "controls_shakenEND.pdf"), width = 10, height = 7)

plot_grid(pA2, pB)+
  ggsave2(here("plots", "controls_shakenBEST.pdf"), width = 10, height = 7)
```


## Calculate RBG
```{r}
rbg.static <-  
d.static%>%
  #choose static with vortexing
  filter(Time!="Tend.vortex")%>%
  group_by(well) %>%
  mutate(Time=ifelse(Time=="T0","start","end"))%>%
  # pivot to join reads of me well
  pivot_wider(id_cols=c("well", "host", "phage"),names_from = Time, values_from=OD600)%>%
  mutate(dif =end-start)%>%
  # divide dif of each phage by no phage
  ungroup()%>%
  group_by(host) %>%
   mutate(rbg.v = dif / dif[phage == 'noPhage'])%>%
  ungroup()%>%
  filter(phage!="noPhage")%>%
  filter(!str_detect(host, "blank"))%>%
    mutate(phage.time= case_when(phage=="futuure" ~ 1,
                               phage=="current" ~ 0,
                               phage=="past" ~ -1))%>%
  mutate(evo.host= if_else(str_detect(host,"ANC"), "anc","evol"))
```


# show time shift curves
```{r}
rbg.static.mean <- rbg.static%>%
  filter(!str_detect(host,"ANC"))%>%
  mutate(rbg.v=if_else(rbg.v>1,1,rbg.v))%>%
  group_by(phage.time)%>%
  summarise(rbg.avg=mean(rbg.v),rbg.sd=sd(rbg.v), .groups="drop" )
  

rbg.static%>%
  filter(evo.host=="evol")%>%
  mutate(rbg.v=if_else(rbg.v>1,1,rbg.v))%>%
  ggplot(aes(x=phage.time, y=rbg.v))+
  geom_line(aes(group=host,color=host), alpha=0.5, size=1)+
  geom_line(data=rbg.static.mean,aes(y=rbg.avg), size=2)+
  geom_pointrange(data=rbg.static.mean,
                  aes(y=rbg.avg,ymin=rbg.avg-rbg.sd,ymax=rbg.avg+rbg.sd),
                  size=1)+
  theme_cowplot()+
  scale_x_continuous(breaks=c(-1,0,1))+
  scale_color_grey()+
  theme(legend.position = "none")
```

<!-- * It seems like there is a phage contamination in a few of the control wells: A7, E7 and C12   -->
<!-- * Well H8 may have not recieved phage. -->
<!-- I will remove those wells:   -->
<!-- ```{r} -->
<!-- contam.wells <- c("A7", "E7" , "C12", "H8") -->
<!-- d <-  -->
<!--   d%>% -->
<!--   filter(!well %in%contam.wells) -->

<!-- meta <-  -->
<!--   meta%>% -->
<!--   filter(!well %in%contam.wells) -->
<!-- ``` -->




<!-- ## Calculating the viruence index  -->
<!-- Folowing Storms et. al.   -->
<!-- *1. establish limit of integration* -->
<!-- *"It is important to stress how the establishment of the limit of integration plays a significant role in the assessment of virulence. This limit should be set as the onset of stationary phase in the phage-free control. This provides a consistent reference for integration that can be easily identified for any phage–host system and restricts measurements to the period of cell growth—a necessary condition for productive infection for many phages.47 Moreover, it ensures that the range of the virulence measurements is well distributed, as discussed hereunder. In general, we recommend establishing the limit of integration as the time at which the slope of OD630 over time reaches ≤0.03 h."* -->

<!-- ### Determining the integration limit -->
<!-- Doing this can be done in 2 ways:   -->
<!-- A.using 'growthcurver' to find no-phage carying capacity (k) -->
<!-- ```{r} -->
<!-- library(growthcurver) -->

<!-- k.noPHI <-  -->
<!--   input[input$Time<8,]%>% -->
<!--   select(c("Time", meta$well[meta$phage=="noPHI"]))%>% -->
<!--   SummarizeGrowthByPlate(., bg_correct = "none")%>% -->
<!--   select("sample","k") -->
<!-- colnames(k.noPHI) <- c("well" ,"k")  -->


<!-- k.noPHI$Time.k <- NA -->
<!-- for(i in seq(k.noPHI$well)){ -->
<!--   w <- k.noPHI$well[i] -->
<!--   k <- k.noPHI$k[i] -->
<!--   tmp <-  -->
<!--     d%>% -->
<!--     filter(well==w) -->
<!--   k.noPHI$Time.k[i] <- -->
<!--     tmp$Time[which(tmp$OD600>k)[1]] -->

<!-- } -->
<!-- rm(tmp, i,w,k) -->
<!-- int.limit <- median(k.noPHI$Time.k, na.rm = TRUE) -->
<!-- int.limit -->
<!-- ``` -->

<!-- B. by visual inspection of plots   -->
<!-- First I will look at the change in OD over time. Storms et al. reccomend *establishing the limit of integration as the time at which the slope of OD630 over time reaches ≤0.03 h*   -->
<!-- ```{r} -->
<!-- time.diff <- diff(d$Time[1:2]) -->

<!-- d %>%  -->
<!--   filter(phage=="noPHI")%>% -->
<!--   split(.$well)%>%  -->
<!--   map_df("OD600")%>% -->
<!--   map_df(diff)%>% -->
<!--     map_df(function(x) x/time.diff)-> tmp -->

<!-- matplot(x= unique(d$Time)[-60],tmp, type="l", ylab = "OD/Time diff") -->
<!-- abline(h=0.03, col="red", lwd=3) -->
<!-- abline(v=c(1:floor(max(d$Time))), col="grey") -->

<!-- ``` -->

<!-- Both methods show that at integrating limit should be at ~ 4 hours. That is what I will set. How does that look?   -->
<!-- ```{r} -->
<!-- int.limit <- 4 -->

<!-- d%>% -->
<!--   filter(phage=="noPHI")%>% -->
<!--   filter(Time<int.limit)%>% -->
<!--   ggplot(aes(x=Time, y=OD600))+ -->
<!--   geom_line(data=filter(d,phage=="noPHI"), color="grey", size=1)+ -->
<!--   geom_line(aes(color=as.factor(dilution)), size=1)+ -->
<!--   # geom_hline(data=k.noPHI, aes(yintercept=k) )+ -->
<!--   facet_wrap(~well)+ -->
<!--   theme_cowplot()+ -->
<!--     theme(legend.position = "none") -->

<!-- ``` -->

<!-- not quite at stationery. Lets try an integration limit of T<5   -->
<!-- ```{r} -->
<!-- int.limit <- 5 -->

<!-- d%>% -->
<!--   filter(phage=="noPHI")%>% -->
<!--   filter(Time<int.limit)%>% -->
<!--   ggplot(aes(x=Time, y=OD600))+ -->
<!--   geom_line(data=filter(d,phage=="noPHI"), color="grey", size=1)+ -->
<!--   geom_line(aes(color=as.factor(dilution)), size=1)+ -->
<!--   # geom_hline(data=k.noPHI, aes(yintercept=k) )+ -->
<!--   facet_wrap(~well)+ -->
<!--   theme_cowplot()+ -->
<!--     theme(legend.position = "none") -->



<!-- ``` -->

<!-- I think that is better.   -->

<!-- ## Choosing the dilution range   -->
<!-- Need to make sure to use only dilutions in which all wells got phage.   -->
<!-- ```{r} -->
<!-- phage.vol <- 0.1 #ml -->
<!-- init.titer <- 1.5e11 #PFU/ml -->
<!-- max.dilut <- 1/(init.titer*phage.vol) -->
<!-- #-# verify visualy -->
<!-- d%>% -->
<!--   filter(host != "Blank")%>% -->
<!--   # filter(Time<int.limit)%>% -->
<!--   filter(phage!="noPHI")%>% -->
<!--   # filter(dilution<=max.dilut)%>% -->
<!--   ggplot(aes(x=Time, y=OD600))+ -->
<!--   geom_line(aes(color=well), size=1)+ -->
<!--   facet_grid(dilution~interaction(host, phage))+ -->
<!--   theme_cowplot()+ -->
<!--     theme(legend.position = "none") -->

<!-- ``` -->

<!-- All the dilutions can be used.  -->

<!-- ```{r} -->
<!-- # d%>% -->
<!-- #   filter(host != "Blank")%>% -->
<!-- #   # filter(Time<int.limit+1)%>% -->
<!-- #   # filter(phage!="noPHI")%>% -->
<!-- #   filter(dilution==1e-7 |dilution==1e-6 |dilution==1e-5)%>% -->
<!-- #   ggplot(aes(x=Time, y=OD600))+ -->
<!-- #   geom_line(aes(color=well), size=1)+ -->
<!-- #   facet_grid(dilution~interaction(host, phage))+ -->
<!-- #   theme_cowplot()+ -->
<!-- #     theme(legend.position = "none") -->
<!-- #  -->
<!-- # max.dilut <-1e-7 -->
<!-- ``` -->

<!-- Plot of lysis -->
<!-- ```{r} -->
<!-- d%>% -->
<!--   filter(host != "Blank")%>% -->
<!--   filter(dilution != "Blank")%>% -->
<!--   filter(Time<10)%>% -->
<!--   filter(!grepl("COL",dilution))%>% -->
<!--   # filter(phage!="noPHI")%>% -->
<!--   # filter(dilution<=max.dilut)%>% -->
<!--   ggplot(aes(x=Time, y=OD600))+ -->
<!--   geom_line(aes(group=well,color=host), size=1)+ -->
<!--   facet_wrap(~fct_rev(dilution))+ -->
<!--   theme_cowplot()+ -->
<!--   ggtitle("SPO1 infection by MOI (n=4)") -->
<!--     # theme(legend.position = "none") -->
<!-- ``` -->

<!-- ## Integrate the area under curve   -->
<!-- Storms et al.: *"areas underneath the optical density versus time curves were calculated using the trapezoid rule for each well, from the time of infection to the  time corresponding to the onset of stationary phase in the phage-free control* -->

<!-- ```{r} -->
<!-- library(caTools) #needed for trapezoids -->
<!-- meta$auc <- NA -->
<!-- for(i in seq(meta$well)){ -->
<!--   if(meta$host[i]=="Blank") next -->

<!--   tmp <-  -->
<!--     d%>% -->
<!--     filter(well==meta$well[i])%>% -->
<!--     filter(Time<=int.limit) -->
<!--   meta$auc[i] <- trapz(tmp$Time, tmp$OD600) -->

<!-- } -->
<!-- rm(tmp) -->

<!-- # summarize no phage control areas -->
<!-- sum.noPHI <-  -->
<!--   meta%>% -->
<!--     filter(host != "Blank")%>% -->
<!--     filter(phage=="noPHI")%>% -->
<!--     group_by(host, colony)%>% -->

<!--     summarise( A0=mean(auc), sd=sd(auc)) -->
<!-- # %>% -->
<!--   # ggplot(aes(colony, mean))+ -->
<!--   # geom_pointrange(aes(ymin=mean-sd, ymax=mean+sd))+ -->
<!--   # facet_wrap(~host) -->
<!-- vindex <- merge(meta, sum.noPHI) -->
<!-- vindex$Vi <- 1-(vindex$auc/vindex$A0) -->
<!-- vindex$log.moi <- log10(as.numeric(vindex$dilution)*1.5e11/2.5e7) -->
<!-- vindex%>% -->
<!--   filter(phage!="noPHI")%>% -->
<!--   group_by(host,colony, phage, log.moi)%>% -->
<!--   summarise( Virulence=mean(Vi), mn=min(Vi), mx=max(Vi), n=n())%>% -->
<!--   ggplot(aes(log.moi, Virulence, color=colony))+ -->
<!--            geom_pointrange(aes(ymin=mn, ymax=mx ), size=1)+ -->
<!--            geom_line(aes(group=colony), size=1)+ -->
<!--            facet_wrap(~interaction(phage, host), nrow=2)+ -->
<!--            theme_cowplot()+ -->
<!--   ggtitle("interaction(host,phage)") -->
<!-- ``` -->

<!-- ## summarize to single value index -->
<!-- Storms et al.: *The virulence index is defined as the area under the virulence curve (AP) divided by the theoretical maximum area under the virulence curve (Amax)* -->
<!-- **Vp = Ap / Amax**   -->
<!-- ```{r} -->
<!-- sum.phi <- vindex%>% -->
<!--   filter(phage!="noPHI")%>% -->
<!--   group_by(host,colony, phage)%>% -->
<!--   summarise(  n=n()) -->

<!-- sum.phi$Ap <- NA -->
<!-- sum.phi$Amax <- NA -->
<!-- for(i in seq(nrow(sum.phi))){ -->

<!--   tmp <-  -->
<!--     vindex%>% -->
<!--     filter(phage!="noPHI")%>% -->
<!--     filter(host==sum.phi$host[i])%>% -->
<!--     filter(colony==sum.phi$colony[i])%>% -->
<!--     filter(phage==sum.phi$phage[i])%>% -->
<!--     # filter(dilution>=max.dilut)%>% -->
<!--     arrange(log.moi) -->
<!--   sum.phi$Ap[i] <- trapz(tmp$log.moi, tmp$Vi) -->
<!--   sum.phi$Amax[i] <-  trapz(tmp$log.moi,rep(1, nrow(tmp))) -->

<!-- } -->

<!-- rm(tmp) -->
<!-- sum.phi$Vp <- sum.phi$Ap/sum.phi$Amax -->

<!-- sum.phi%>% -->
<!--   # filter(n==8)%>% -->
<!--   ggplot(aes(x=interaction(phage,host), y=Vp))+ -->
<!--   geom_boxplot(width=0.2)+ -->
<!--     geom_jitter(shape=21,#aes(shape=as.factor(colony)), -->
<!--               height=0, width=0.05,size=5, fill="grey", alpha=0.5)+ -->
<!--   # scale_shape_manual(values=c(15,16,22,21))+ -->
<!--   ylim(0,1)+ -->
<!--   theme_cowplot() -->
<!-- ``` -->

<!-- ## Stats   -->
<!-- ```{r} -->
<!-- summary(aov(Vp~host+colony, sum.phi  )) -->
<!-- ``` -->

<!-- # No Difference!!   -->


<!-- Nice plot -->
<!-- ```{r} -->
<!-- d%>% -->
<!--   filter(dilution != "Blank")%>% -->
<!--   filter(Time<10)%>% -->
<!--   filter(phage!="noPHI")%>% -->
<!--   # filter(dilution<=max.dilut)%>% -->
<!--   ggplot(aes(x=Time, y=OD600))+ -->
<!--   geom_line(aes(group=well,color=dilution), size=1)+ -->
<!--   facet_grid(host~colony)+ -->
<!--   theme_cowplot()+ -->
<!--     theme(legend.position = "bottom") -->
<!-- ``` -->

